## Sample Project using survey data

```{r}
library(readxl)
sample_data_1<- read_excel("C:/Users/hp/Downloads/sample_data_1_1_.xlsx")
```

## Exploring the data
```{r}
library(tidyverse)
library(dplyr)
library(haven)
```
Using glimpse form the dplyr package we are able to see how our data looks like.Also using head() and tail() e can see the first five rows and last 5 rows.

```{r}
glimpse(sample_data_1)
head(sample_data_1,5)
tail(sample_data_1,5)
```

Obtaining the summary to get the picture of our data

```{r}
summary(sample_data_1)
```

## Sample sizes for education subgroup

```{r}
sample_data_1 %>%
  group_by(v149) %>%
  summarise(n())
```

## Sample sizes and proportion of records for education subgroup

```{r}
sample_data_1 %>%
  group_by(v149) %>%
  summarise(n = n()) %>%
  mutate(p = n/sum(n))
```

## Constructing health indicators (skilled birth attendance)
```{r}
sample_data_1a <- sample_data_1 %>%
  mutate(sba =
           if_else(m3a == "yes"|
                   m3b == "yes"|
                   m3c == "yes"|
                   m3d == "yes",
                   1,0)
         )
```

# Tabulate *sba* variable (skilled birth attendance)
```{r}
count(sample_data_1a, sba)
```

# # Replacing missing values in *sba* with zeros and inspecting results
```{r}
sample_data_1b <- sample_data_1a %>%
  mutate(sba =
           if_else(m3a == "yes"|
                   m3b == "yes"|
                   m3c == "yes"|
                   m3d == "yes",
                   1,0,0)
         )
count(sample_data_1b, sba)
```

# Inspecting R objects
```{r}
sample_data_1b
head(sample_data_1b$sba , n = 10)
```


